A Comparison of Graph Processing Systems

### 2020-09-19 documentation/paper/main.tex
some work on discussion/comparison

\usepackage{tabularx}
\def\tabularxcolumn#1{m{#1}} % vertical centering in tabularx cells
\newcolumntype{Y}{>{\centering\arraybackslash}X} % cell type Y for centered text with automatic line breaks
\newcommand*\rot{\rotatebox{45}} % rotates content 90deg ccw, mainly for table headers
\newcommand*\OK{\ding{51}} % a checkmark-style sign

@@

\section{Discussion}
\todo{Hier zunächst mal ein kurzer Vergleich zwischen Distributed und Single Node wo sinnvoll, dann zwischen den Frameworks und so weiter..}

>>>>

\usepackage{tabularx}
\def\tabularxcolumn#1{m{#1}} % vertical centering in tabularx cells
\newcolumntype{Y}{>{\centering\arraybackslash}X} % cell type Y for centered text with automatic line breaks
\newcommand*\rot{\rotatebox{45}} % rotates content 90deg ccw, mainly for table headers
\newcommand*\OK{\ding{51}} % a checkmark-style sign
\newcommand{\txtdagger}{\textsuperscript{\textdagger{}}}

@@

\section{Discussion}
\todo{unfertig}

There are two main cases that should be regarded in a comparison of the frameworks. One, the case of individual calculations on a graph, i.e. for each calculation, the graph has to be loaded. For this case, frameworks with small execution times and small overhead are preferred. Case two on the other hand is the case of a running system, that performs multiple calculations on a single graph, without the need of reloading graph data with every calculation.
In this case, frameworks with short calculation times should be preferred because the overhead time is only spent once on startup.

Orthogonal to this, the graph size and topology of the system is a decinding factor. Large graphs often require a distributed cluster, because the graph is simply too large to fit in the RAM of just one machine. We observed this for example with Giraph on 4 of 7 graphs. Furthermore, some environments require a distributed framework setup regardless of any other factors.

SSSP
For graphs that are not too large for the RAM and individual test runs, Galois as the fastest single-node framework should be utilized. It provides excellent performance, being orders of magnitude faster than the competing frameworks on a single node.

Especially for frameworks like Giraph, that require large amounts of memory, the distributed scenario allows them to complete their computations.

BFS

PR

### 2020-09-19 documentation/paper/sections/relatedWork.tex
some work on discussion/comparison

\section{Related Work}
Almost every framework, that gets published in a paper compares itself to other, similar frameworks. 
Hence there already exists some information on the relative performance between frameworks. 
However, many of the most common frameworks today have been released many years ago. 
Because of that two problems arise.
First, the comparisons are often themselves very old. 
And second, many other frameworks are no longer in use because they were replaced by something else.
We are thus providing an objective benchmark on what changed since the initial release, with a comparison between the most commonly used frameworks today.

The most recent publication on Galois was in 2013 \cite{Galois}, along with a comparison to GraphLab, PowerGraph and Ligra, which is in our testing lineup itself. The results of their comparison is mainly, that Galois outperforms Ligra on BFS and SSSP on most graphs. Meanwhile, their performance on PR is similar but very dependent on the graph. In some cases, Ligra outperforms Galois, on other graphs Galois is faster than Ligra.

Gemini puts itself against Ligra, Galois, PowerGraph, PowerLyra and GraphX \cite{Gemini}.
In a single-threaded scenario, Ligra is faster than both Galois and Gemini in SSSP and BFS applications. 
On the other hand, Gemini beats the other two in PR.
Furthermore, Gemini is presented to be one order of magnitude faster than PowerGraph, PowerLyra and GraphX on a multi-node computation unit.

>>>>

\section{Related Work}
\todo{Der erste Abschnitt hier kann eigentlich so wie er ist auch in die Introduction.}

Almost every framework, that gets published in a paper compares itself to other, similar frameworks. 
Hence there already exists some information on the relative performance between frameworks. 
However, many of the most common frameworks today have been released many years ago. 
Because of that, two problems arise.
First, the comparisons are based on very old versions of the frameworks, with no information on the current behaviour available. 
And second, many of the other frameworks are no longer in use because they were replaced by something else, making the comparisons difficult to interpret.
We are thus providing an objective benchmark on what changed since the initial release, with a comparison between the most commonly used frameworks today.

It quickly becomes apparent that many of the comparisons provided

The comparisons often include PowerGraph or Apache's Spark GraphX, two frameworks that we did not include in our testing.

The most recent publication on Galois was in 2013 \cite{Galois}, along with a comparison to GraphLab, PowerGraph and Ligra, which is in our testing lineup itself. The results of their comparison is mainly, that Galois outperforms Ligra on BFS and SSSP on most graphs. Meanwhile, performance of Ligra and Galois on PR is similar but very dependent on the graph. In some cases, Ligra outperforms Galois, on other graphs Galois is faster than Ligra.

Gemini puts itself against Ligra, Galois, PowerGraph, PowerLyra and GraphX \cite{Gemini}.
In a single-threaded scenario, Ligra was shown to be faster than both Galois and Gemini in SSSP and BFS applications. 
On the other hand, Gemini beat the other two in PR applications.
Furthermore, Gemini is presented to be one order of magnitude faster than PowerGraph, PowerLyra and GraphX on a multi-node computation unit.

### 2020-09-19 documentation/paper/sections/results/results.tex
some work on discussion/comparison

>>>>

\subsubsection{Comparison of a Single-Node to the Distributed Cluster}
\begin{table}
	\caption{Execution Times on a SingleNode vs. 5-Node Distributed Cluster}
	\label{tbl:SSSP_execTimeComparison}
	\renewcommand{\arraystretch}{1.2}
	\centering
	\begin{tabular}{ccr@{\tabskip 1 \tabcolsep}r
	r@{\tabskip 1 \tabcolsep}r
	r@{\tabskip 1 \tabcolsep}r}
		\toprule
		&&\multicolumn{2}{c}{\bf Galois}&\multicolumn{2}{c}{\bf Gemini}&\multicolumn{2}{c}{\bf Giraph}\\
		\cmidrule{3-4}\cmidrule{5-6}\cmidrule{7-8}
		&\bf Graph&1N&5N&1N&5N&1N&5N\\
		\midrule
		\multirow{7}{0.5ex}{\rotatebox{90}{\bf SSSP}}&flickr & 0.3 & 2.4 & 0.6 & 2.5 & 38.8 & 42.6 \\
		& orkut & 0.8 & 4.2 & 13.9 & 17.3 & 135.9 & 60.4 \\
		& wikipedia & 1.8 & 14.2 & 48.5 & 74.5 & 377.2 & 100.8 \\
		& twitter & 10.8 & 40.0 & 383.0 & 391.1 & - & 349.9 \\
		& rMat27 & 16.0 & 39.0 & 563.5 & 386.0 & - & 565.8 \\
		& friendster & 14.4 & 58.5 & 742.6 & 568.9 & - & 444.0 \\
		& rMat28 & 27.8 & 71.5 & 1236.7 & 792.0 & - & 1180.2 \\
		\midrule
		\multirow{7}{0.5ex}{\rotatebox{90}{\bf BFS}}& flickr & 0.6 & 2.3 & 0.7 & 2.4 & 38.9 & 41.9 \\
		& orkut & 0.9 & 4.0 & 12.5 & 13.8 & 128.2 & 58.1 \\
		& wikipedia & 2.4 & 13.2 & 44.9 & 60.9 & 360.7 & 95.1 \\
		& twitter & 14.2 & 37.9 & 355.0 & 293.2 & - & 322.3 \\
		& rMat27 & 14.6 & 34.4 & 540.8 & 305.3 & - & 539.2 \\
		& friendster & 12.8 & 47.6 & 708.4 & 450.2 & - & 412.3 \\
		& rMat28 & 33.1 & 67.1 & 1178.7 & 634.3 & - & 1135.6 \\
		\midrule
		\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR}}& flickr & 0.3\txtdagger & 2.6 & 0.7 & 2.2 & 45.6 & 44.1 \\
		& orkut & 0.7\txtdagger & 8.0 & 13.1 & 19.4 & 561.3 & 106.2 \\
		& wikipedia & 1.7\txtdagger & 206.5 & 44.4 & 54.4 & 1501.3 & 252.6 \\
		& twitter & 8.7\txtdagger & 910.9 & 359.9 & 363.5 & - & 1200.2 \\
		& rMat27 & 19.2\txtdagger & 1287.4 & 536.3 & 376.3 & - & 1369.5 \\
		& friendster & 20.4\txtdagger & 593.3 & 716.5 & 591.1 & - & 1655.0 \\
		& rMat28 & 46.0\txtdagger & 2540.2 & 1188.5 & 774.2 & - & - \\
		\bottomrule
		\multicolumn{8}{l}{(\txtdagger) Results of Galois Pull shown.}
	\end{tabular}
\end{table}
Until now, we have only compared the frameworks with each other, under the same setup circumstances. Hence, a comparison of the same framework, in single-node versus distributed setup is the content of this section.
Of course only frameworks that were tested in both setups are shown.
We focus primarily on the results of the execution times, the data can be seen in \autoref{tbl:SSSP_execTimeComparison}. We show the results of Galois Push for the 5-node cluster in the table, since it is faster than Pull.

There are some, rather large differences in the execution times of each framework. Galois is consistently faster on a single computation node compared to the distributed setup. 
On both SSSP and BFS, the margin between the two setups is already noticeable. The distributed scenario requires from 2$\times$ (BFS, rMat28) to 7.8$\times$ (SSSP, wikipedia) the execution time of single-node Galois.
For PR however, the difference is multiple orders of magnitude large. Distributed Galois requires from 8.6$\times$ to 121$\times$ more time than single-node Galois.

Gemini's distributed calculation only requires more time on the smaller graphs (i.e. flickr to wikipedia). On the larger graphs, the added computation power can be leveraged to overcome the synchronization overhead of the distributed scenario. On rMat27, friendster and rMat28, Gemini is anywhere from 1.3$\times$ to 1.6$\times$ faster than on a single-node.
Giraph already uses the distributed cluster efficiently on the smaller graphs as well. Only on flickr is the distributed version slower than a single node. The differences for Giraph on the other graphs are rather large, single-node execution time is 2.25$\times$ and 3.7$\times$ the execution time of the distributed version on orkut and wikipedia respectively.
